---
title: "Assignment 9"
output: pdf_document
---
## Question 12.1

Describe a situation or problem from your job, everyday life, current events, 
etc., for which a design of experiments approach would be appropriate.

## Answer 12.1

I am presently working towards loosing weight. This is a very good situtation 
where I can use DOE to see what factors (execercise and diet with their 
variations) will help or have the largest impact on my targets. It can also help 
in identifying important interactions between the factors. 


## Question 12.2

To determine the value of 10 different yes/no features to the market value of a 
house (large yard, solar roof, etc.), a real estate agent plans to survey 50 
potential buyers, showing a fictitious house with different combinations of 
features.  To reduce the survey size, the agent wants to show just 16 fictitious 
houses. Use R’s FrF2 function (in the FrF2 package) to find a fractional 
factorial design for this experiment: what set of features should each of the 16 
fictitious houses have?  Note: the output of FrF2 is “1” (include) or  “-1” 
(don’t include) for each feature.

## Answer 12.2

2. Experimental Design
With the current selection we have 10 factors and 16 runs.Factorials designs are 
useful as screening experiments because they require relatively few runs to 
estimate main and interaction effects.

Randomization
While we have no control over how the data will be collected, by selecting a 
random sample from the data, we are incorporating randomization into the model.

Blocking
Blocking is not used in this design. This is a screening experiment and we are 
interested in the effects of all the factors.

```{r message=FALSE, warning=FALSE}
library(FrF2)
```

```{r message=TRUE, warning=FALSE}
set.seed(42)
des<-FrF2(nfactors = 10,nruns = 16, 
          factor.names = c('Yard', 'Parking','Rooms','Basement', 'Price',
                           'Location','School Nearby', 'Supermarket Nearby',
                           'Luxary Home','Train Station Nearby'))
```

## Question 13.1

For each of the following distributions, give an example of data that you would 
expect to follow this distribution (besides the examples already discussed 
in class).

a.	Binomial   
b.	Geometric   
c.	Poisson   
d.	Exponential  
e.	Weibull


## Answer 13.1

A probability distribution is a statistical function that describes all the 
possible values and likelihoods that a random variable can take within a 
given range. This range will be bounded between the minimum and maximum possible 
values, but precisely where the possible value is likely to be plotted on 
the probability distribution depends on a number of factors. These factors 
include the distribution's mean (average), standard deviation, skewness, and 
kurtosis.

#	Binomial

The binomial distribution describes the behavior of a count variable X if 
the following conditions apply:

1: The number of observations n is fixed.
2: Each observation is independent.
3: Each observation represents one of two outcomes ("success" or "failure").
4: The probability of "success" p is the same for each outcome.

If these conditions are met, then X has a binomial distribution with 
parameters n and p

Example: Applying for a n Data Scientist jobs with observation being selected or 
rejected. Thus, the number of times a person gets selected might follow 
a binomial distribution

#	Geometric   

If I randomly ask people which university they went to, until I meet someone who 
went to Georgia Tech. Thus, in this example p is the probaility of finding 
someone who went to Georgia Tech after asking X number of people. In this case, 
we can say that X follows a geometric distribution.

# Poisson   
The number of car accidents per hour in a day might follow a Poisson Distribution

# Exponential  
In the above example the time between two accidents might follow the exponential 
distribution. 

# Weibull
Modeling the time to failure of an engine by a car manufacturer might follow
Weibull distribution.

## Question 13.2

In this problem you, can simulate a simplified airport security system at a busy 
airport. Passengers arrive according to a Poisson distribution with λ1 = 5 per 
minute (i.e., mean interarrival rate mu1 = 0.2 minutes) to the ID/boarding-pass 
check queue, where there are several servers who each have exponential service 
time with mean rate mu2 = 0.75 minutes. [Hint: model them as one block that has 
more than one resource.]  After that, the passengers are assigned to the shortest 
of the several personal-check queues, where they go through the personal scanner
(time is uniformly distributed between 0.5 minutes and 1 minute). 

Use the Arena software (PC users) or Python with SimPy (PC or Mac users) to build 
a simulation of the system, and then vary the number of ID/boarding-pass checkers
and personal-check queues to determine how many are needed to keep average wait 
times below 15 minutes.  [If you’re using SimPy, or if you have access to a non-
student version of Arena, you can use λ1 = 50 to simulate a busier airport.]

## Answer 13.2



```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
